---
title: "Modeling Healthcare Facility Functionality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling Healthcare Facility Functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `vrcmort` package allows incorporating the impact of healthcare facility functionality on mortality and reporting rates using a strictly monotonic structure. This feature is particularly useful when you have data on the number of facilities at different levels of functionality (e.g., from HeRAMS or similar assessments).

## Monotonic Constraints

We assume that facility levels are provided in **decreasing order of functionality**:
1. **Level 1**: Perfectly functioning.
2. **Level 2**: Partially functioning.
3. ...
4. **Level K**: Minimally functioning.

The model assumes a baseline of **non-functioning facilities** (all provided levels = 0) which has zero effect on the log-mortality or logit-reporting.

### Mortality ($\lambda$)
Better functioning facilities are expected to reduce mortality more than less functional ones. Therefore, the coefficients $\beta$ follow:
$$\beta_1 < \beta_2 < \dots < \beta_K < 0$$

### Reporting ($
ho$)
Better functioning facilities are expected to increase reporting rates more than less functional ones. Therefore, the coefficients $\gamma$ follow:
$$\gamma_1 > \gamma_2 > \dots > \gamma_K > 0$$

## Usage

To use this feature, provide the names of the columns representing the facility counts to the `facilities` argument in `vrc_mortality()` or `vrc_reporting()`.

```{r example, eval = FALSE}
library(vrcmort)

# Assume df has columns 'perfect_fac', 'partial_fac', 'minimal_fac'
# representing standardized counts or proportions.

fit <- vrcm(
  mortality = vrc_mortality(
    facilities = c("perfect_fac", "partial_fac", "minimal_fac")
  ),
  reporting = vrc_reporting(
    facilities = c("perfect_fac", "partial_fac")
  ),
  data = df,
  t0 = "2023-01-01"
)

# Summarise the inferred coefficients
vrc_coef_summary(fit)
```

## Identifiability

The model uses a "Gaps" parameterization to ensure identifiability and efficient sampling. It estimates:
- The effect of the "best" level (Level 1).
- A simplex of "gap ratios" that distribute this total effect across the intermediate levels down to the zero baseline.

This structure guarantees that the strictly monotonic relationships are maintained throughout the posterior.
